---
title: "Gary's analysis"
author: "Gary Sztajnman"
date: "February 16, 2016"
output: pdf_document
geometry: margin=2cm
---

# Part 4 - Understanding the impact of flood

Flood have a huge impact on population and economy.
Heavy rain leads to destructionsm, death, displacement, healness, reparations.
Some countries may receive more floods based on its location (for example India) but some countries also don't have the capacity to prevent, prepare population, manage flood and repare.

In this analysis we will try to analyse what characteristic influence the impact of a flood.

## Impact evolution

```{r, echo=FALSE, warning= FALSE}
library(scales)
library(chron)
library(RColorBrewer)
library(ggplot2)


workdir ="/Users/garyair/Desktop/Dropbox/Columbia/EDA/P2/Edav-project-2/Edav-project-2/Data"
setwd(workdir)


Dev = read.csv("2015_development index.csv")
#View(Dev)
flood2 <- read.csv("cleanGlobalFloodDataWithRegionsfiltered-Index noNA.csv")
#View(flood2)

## Adding a year column
#date1<- as.Date(flood$dateBegan,"%m/%d/%Y")
#flood$year<-as.numeric(format(date1, "%Y"))

## Add other to mainCause1

flood2$GNIPerCapita <- as.numeric(flood2$GNIPerCapita)
flood2$peopleDisplaced <- as.numeric(flood2$peopleDisplaced)
flood2$HDIclass <- cut(flood2$HDI,
                     breaks=c(-Inf, 0.6, 0.75, Inf),
                     labels=c("low","medium","high"))
#summary(flood2$HDIclass)
#plot(flood2$HDIclass)

```

Let's first take a look at the evolution of flood in time

```{r, echo=FALSE}
#PLOT
## Severity by year:
ggplot(data = flood2, aes(year, magnitude, fill=severity)) +
  stat_summary(fun.y = "mean", geom = "bar")  # or "line" 

## Graph by cause
#ggplot(data = flood2, aes(year, magnitude, fill=mainCause1)) +
#  stat_summary(fun.y = "sum", geom = "bar")  # or "line" 

## Graph by cause except Heavy rain
#ggplot(subset(flood2, mainCause1 != "Heavy Rain")) +
#  aes(year, magnitude, fill=mainCause1) +
#  stat_summary(fun.y = "sum", geom = "bar")
```

We note that there is an augmentation of the severity of flood in the last decade. But there is no specific cause that can explain it.

```{r, echo=FALSE}
## Type of flood by continent
ggplot(subset(flood2)) +
  aes(year, magnitude, fill=mainCause1) +
  stat_summary(fun.y = "sum", geom = "bar") +
  facet_wrap(~sub_region, ncol = 4)
```

When we analyze the repartion of floods by region, we note that there are some region more affected and that some region are more targeted by specific type of disaster.
For example:

 + Eastern Europe and North American are more touched by Ice Melt
 + Southern Asia is highly affected by Monsoon.
 + Central America is mainly touched by hurricane.
  

# Impact on population
```{r, echo=FALSE}
# Multiple plot function
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

## Type of flood by continent
p1 <- ggplot(subset(flood2)) +
  aes(year, peopleDead, fill=mainCause1) +
  stat_summary(fun.y = "sum", geom = "bar") 

p2 <- ggplot(subset(flood2)) +
  aes(year, peopleDisplaced, fill=mainCause1) +
  stat_summary(fun.y = "sum", geom = "bar")+
   guides(fill = FALSE)

multiplot(p1, p2, cols=2)
```

Based on these 2 graphs, we note 2 things:

 + Some floods have terrible impact on human lives. For example, in Thailand, in 2004 when 160,000 people died in a tsunami. Therefore, we assume that the number of dead is not linked to the characteristic of the country.
 + When we look at the number of people displaced, we see a very different patern. It is quite stable and the main event that push people to move is Monsoon. Furthermore, we know that monsoon tend to happend in a very specific region of the world, mainly south east Asia.
 

## Characteristics of country and floods' impact

In what type of countries does flood have a greater impact on population?
To answer this question we will gather and merge data about Human development index (HDI), life expectancy, expected number of school year, Gross National Income (GNI) per capita.

```{r, echo=FALSE}
## People displaced/magnitude of flood by year, HDIclass and sub-continent
ggplot(subset(flood2, mainCause1 = "Dam failure")) +
  aes(year, peopleDisplaced, fill= HDIclass) +
  stat_summary(fun.y = "sum", geom = "bar") +
  facet_wrap(~region, ncol = 3)
```

When we look at people displaced, we see that it mainly affect Asian and African countries and mainly countries with low (below 0.6) or medium (between 0.6 and 0.75) HDI.

## Regression Model



```{r, echo=FALSE}
summary(lm(peopleDead ~ HDI + lifeExp+ GNIPerCapita, data = flood2))$coefficients


```

The regression analysis confirm that country characteristics does not impact the number of dead people per flood. 


```{r, echo=FALSE}
fit1 =  lm(peopleDisplaced ~ HDI + lifeExp+ GNIPerCapita, data = flood2)
summary(fit1)
```

There is a true correlation between the stage of development of a country and the number of displaced people during floods.

## Main finding

When it comes to death toll, no country is protected against a huge event as a huge tsunami or a hurricane. But facing heavy rain, developed country have better infrastructure and a greater capacity to take care of the people touched by such event. They also have the ability to quickly repair in order to make the population suffer a minimum time so that they don't have to move.




