---
title: "Gary's analysis"
author: "Gary Sztajnman"
date: "February 16, 2016"
output: pdf_document
---

/section{Hypothesis}

Flood have a huge impact on population and economy.
Heavy rain leads to destructionsm, death, displacement, healness, reparations.
Some countries may receive more floods based on its location (for example India) but some countries also don't have the capacity to prevent, prepare population, manage flood and repare.

In this analysis we will try to analyse what characteristic influence the impact of a flood.



```{r, echo=FALSE}
library(gdata)
library(scales)
library(chron)
library(RColorBrewer)
library(lattice)
library(ncdf4)
library(maps)
library(ggplot2)
library(ggmap)
library(mapdata)

workdir ="/Users/garyair/Desktop/Dropbox/Columbia/EDA/P2/Edav-project-2/Edav-project-2/Data"
setwd(workdir)

flood = read.csv("cleanGlobalFloodData.csv")
View(flood)

Dev = read.csv("2015_development index.csv")
View(Dev)
flood2 <- read.csv("cleanGlobalFloodDataWithRegionsfiltered-Index noNA.csv")
View(flood2)

## Adding a year column
#date1<- as.Date(flood$dateBegan,"%m/%d/%Y")
#flood$year<-as.numeric(format(date1, "%Y"))

## Add other to mainCause1

flood2$GNIPerCapita <- as.numeric(flood2$GNIPerCapita)
flood2$HDIclass <- cut(flood2$HDI,
                     breaks=c(-Inf, 0.6, 0.75, Inf),
                     labels=c("low","medium","high"))
#summary(flood2$HDIclass)
#plot(flood2$HDIclass)

```

Let's first take a look at the evolution of flood in time
```{r, echo=FALSE}
#PLOT
## Severity by year:
ggplot(data = flood2, aes(year, magnitude, fill=severity)) +
  stat_summary(fun.y = "mean", geom = "bar")  # or "line" 

## Graph by cause
#ggplot(data = flood2, aes(year, magnitude, fill=mainCause1)) +
#  stat_summary(fun.y = "sum", geom = "bar")  # or "line" 

## Graph by cause except Heavy rain
#ggplot(subset(flood2, mainCause1 != "Heavy Rain")) +
#  aes(year, magnitude, fill=mainCause1) +
#  stat_summary(fun.y = "sum", geom = "bar")
```

We note that there is an augmentation of the severity of flood in the last decade. But there is no specific cause that can explain it.

```{r, echo=FALSE}
## Type of flood by continent
ggplot(subset(flood2)) +
  aes(year, magnitude, fill=mainCause1) +
  stat_summary(fun.y = "sum", geom = "bar") +
  facet_wrap(~sub_region, ncol = 4)
```

When we analyze the repartion of floods by region, we note that there are some region more affected and that some region are more targeted by specific type of disaster.
For example:
  - Eastern Europe and North American are more touched by Ice Melt
  - Southern Asia is highly affected by Monsoon.
  - Central America is mainly touched by hurricane.
  
  
/section{Impact of population}


```{r, echo=FALSE, eval=False}
## Type of flood by continent
ggplot(subset(flood2)) +
  aes(year, magnitude, fill=mainCause1) +
  stat_summary(fun.y = "sum", geom = "bar") +
  facet_wrap(~region, ncol = 3)


## Type of flood by continent
ggplot(subset(flood2)) +
  aes(year, magnitude, fill=HDIclass) +
  stat_summary(fun.y = "sum", geom = "bar") +
  facet_wrap(~sub_region, ncol = 3)


## Magnitude of flood by year, HDIclass and continent
ggplot(subset(flood2)) +
  aes(year, peopleDisplaced, fill=HDIclass) +
  stat_summary(fun.y = "sum", geom = "bar") +
  facet_wrap(~region, ncol = 3)

## People Dead of flood by year, HDIclass and continent
ggplot(subset(flood2)) +
  aes(year, peopleDead, fill=HDIclass) +
  stat_summary(fun.y = "sum", geom = "bar") +
  facet_wrap(~region, ncol = 3)


## People displaced/magnitude of flood by year, HDIclass and sub-continent
ggplot(subset(flood2)) +
  aes(year, peopleDisplaced/magnitude, fill=HDIclass) +
  stat_summary(fun.y = "sum", geom = "bar") +
  facet_wrap(~sub_region, ncol = 3)


## People displaced/magnitude of flood by year, HDIclass and sub-continent
ggplot(subset(flood2, mainCause1 = "Dam failure")) +
  aes(year, peopleDisplaced, fill=severity) +
  stat_summary(fun.y = "sum", geom = "bar") +
  facet_wrap(~region, ncol = 3)

summary(flood2$peopleDisplaced)

#PREDICTION
## Trying to predict magnitude
summary(lm(magnitude ~ mainCause1, data = flood2))

## Trying to predict peopleDead
summary(lm(peopleDead ~ mainCause1, data = flood2))

summary(lm(peopleDead ~ magnitude + durationDays, data = flood2))

summary(lm(peopleDead ~ HDI + lifeExp + yearsSchool + GNIPerCapita, data = flood2))

summary(lm(peopleDead ~ GNIPerCapita, data = flood2))

summary(lm(peopleDead ~ lifeExp, data = flood2))

### Only model significant
summary(lm(peopleDead ~ HDI + lifeExp, data = flood2))

#PREDICTION
#People displaced
summary(lm(peopleDisplaced ~ HDI + lifeExp, data = flood2))


worldmap <- map_data ("world")
worldmap$long.recenter <- ifelse(worldmap$long < center - 180 , worldmap$long + 360, worldmap$long)
View(worldmap)
ggmap(worldmap)

m <- get_map("worldmap")
ggmap(m)

map('world')
?get_map

ggplot() + borders("world", colour = "black") + geom_point(data = flood2, aes(centroidX, centroidY, ))

```
